---
title: "Cake Recipe Community Health"
author: "Bryan White"
date: '2017-07-12'
output: html_document
---

This is an analysis of the volume and authorship of cake recipes published on Food52's community 
dirven recipe database.  Covered in this post are general trends on the amount of recipes published,
who's publishing the reicpes and what patterns exist in the communtiy.  It was discovered that while
the rate of new authorship and the activity of frequent authors has declined since 2014, the amount
published per author has increased to keep the total number of new recipes added stable.  

```{r, echo = FALSE, message=F, warning=F}
library(plyr)
library(tidyverse)
library(magrittr)
library(stringr)
library(plotly)
library(NLP)
library(openNLP)
library(lubridate)
library(broom)



recipe_info <- readRDS("recipe_info.RDS")

recipe_info %<>%
  mutate(favorite_count = str_replace_all(favorite_count,",","")) %>% 
  mutate(favorite_count = as.numeric(favorite_count)) %>% 
  mutate(comments_count = as.numeric(comments_count)) %>% 
  filter(!is.na(favorite_count)) 

recipe_info %<>% 
  mutate(recipe_name = str_replace(url, "(\\/recipes\\/[0-9]+\\-)","")) %>% 
  mutate(recipe_name = str_replace_all(recipe_name, "-"," "))
```

# Cumulative number of recipes published over time
Over time, the annual average rate of cake recipes added per day has ranged between 1 recipe every 
2 days (.5 per day) and 6 recipes every 5 days (1.2 recipes per day).  2014 was the period of 
highest growth in the collection of recipes with an average of 1.2 recipes per day.  The plot below 
highlights this period in yellow.  

The plot below also shows that currently the number of cake recipes added per day is cooling down 
slowly from its high in 2014.  The year to date average number of recipes added per day is just over
1 per day.  This average rate of increase still allows for great diversity and breadth of new cakes
to be added to the community collection of recipes, though further slowing of the rate may indicate
the community is less engaged with contributing new material.
```{r, echo = F}
date_summary <- recipe_info %>% 
  group_by(date_published) %>% 
  summarise(count_published = n()) %>% 
  ungroup() 

all_dates <- data_frame(date = seq.Date(min(date_summary$date_published),
                                        max(date_summary$date_published),
                                        by = "day")
                        )
date_summary_full <- left_join(all_dates, date_summary, by = c("date" = "date_published")) 

date_summary_full[is.na(date_summary_full)] <- 0

date_summary_full %<>% 
  mutate(cumulative_published = cumsum(count_published))

# plot_ly(data = date_summary_full, y = ~cumulative_published, x = ~date,
#         type = 'scatter', mode = 'markers' )

date_summary_full %<>% 
  mutate(trend = row_number()) %>% 
  mutate(pub_year = year(date))

date_summary_full_split <- split(date_summary_full,date_summary_full$pub_year)
trendy <- function(split_data){
  # split_data <- test_split[[1]]
lm_trend <- lm(cumulative_published~trend,data = split_data)
split_data$predicted_values <- predict(lm_trend, newdata = split_data)
split_data$rate <- lm_trend$coefficients["trend"]
split_data
}

date_summary_full <- map(date_summary_full_split, trendy) %>% bind_rows()
date_summary_full %<>% filter(predicted_values > 0) %>% rename(Rate = rate)
plot_ly(data = date_summary_full, y=~predicted_values, x = ~date, 
        type = 'scatter', mode = 'markers', color = ~Rate, showlegend=FALSE) %>% 
  layout(
    xaxis = list(range = c(min(date_summary_full$date)-5,max(date_summary_full$date) + 5),
                 title = ""),
    yaxis = list(title = "Number of Recipes"))
```


#Distinct Authors over time and their contributions

During late 2013-2014, the number of distinct recipe authors per month increased to their highest 
levels. Since this period, the number of distinct authors has slowly declined to an average of 19 
authors per month in 2017. 

Interestingly, this decrease in the number of distinct authors has been
compensated by the rapid growth in contributions per author.  In the plot below the increase in 
contributrions per author is noted by color, more red indicates higher average contributions per
author.  In 2017, the average monthly contributions per author is 7, this is a 75% increase in 
authorship since 2015-2016 and a 130% increase from 2013-2014.  This trend is driven primarily by 
professional chefs ('Oh Sweet Day!', 'Jordan Rondel, The Caker',Alice Medrich) and Food52 staff
(Sarah Jampel, Food52).

```{r,echo=F}
distinct_authors <- recipe_info %>% 
  mutate(pub_year = year(date_published)) %>%
  mutate(pub_month = month(date_published)) %>% 
  group_by(pub_year, recipe_author) %>% 
  mutate(per_author_ann_contributions = n()) %>% 
  group_by(pub_year) %>% 
  mutate(avg_ann_contributions = mean(per_author_ann_contributions)) %>% 
  group_by(pub_year,pub_month) %>% 
  summarise(distinct_authors = n_distinct(recipe_author),
            avg_ann_contributions = mean(avg_ann_contributions)) %>%
  ungroup() %>% 
  mutate(year_month = paste0(pub_year,"_",pub_month)) %>% 
  arrange(pub_year,pub_month) %>% 
  mutate(y_m_order = row_number()) %>% 
  mutate(pub_year = as.character(pub_year))

numeric_color_ramp <- function(vals, start_color, end_color){
  vals <- vals %>% unique()
  ii <- cut(vals, breaks = seq(min(vals), max(vals), len = length(vals)*2), 
            include.lowest = TRUE)
  ## Use bin indices, ii, to select color from vector of n-1 equally spaced colors
  colorRampPalette(c("#0868ac", "#e31a1c"))(length(vals)*2)[ii]
}

distinct_authors_color_scale <- distinct_authors %>% 
  group_by(pub_year) %>% 
  summarise(avg_ann_contributions = mean(avg_ann_contributions)) %>% 
  ungroup() %>% 
  arrange(avg_ann_contributions) %>% 
  mutate(color_col = numeric_color_ramp(avg_ann_contributions))
distinct_authors = inner_join(distinct_authors,distinct_authors_color_scale,
                          by = c("pub_year", "avg_ann_contributions")) %>% 
  mutate(year_month = as.Date(paste0(year_month,"_01"),f = "%Y_%m_%d"))
plot_ly(distinct_authors,
        x = ~year_month,
        y = ~distinct_authors, type = 'scatter', mode = 'lines',color = ~pub_year,
        colors = distinct_authors_color_scale$color_col,showlegend=FALSE) %>% 
  layout(
    xaxis = list(range = c(min(date_summary_full$date)-5,max(date_summary_full$date) + 5),
                 title = ""),
    yaxis = list(title = "Distinct Number of Authors"))
```
*More red lines indicate higher rates of recipes contributed per recipe author*
 
#New Authors
The rate of new authors contributing to Food52 has increased so far in 2017 relative to the slow 
down in new author contributions frm 2015-2016.  Combining this plot with the previous plots show 
new authors were the primary contributors to the recipe boom of 2014.

```{r, echo = F}
new_postings <- recipe_info %>% 
  group_by(recipe_author) %>% 
  filter(date_published == min(date_published)) 

new_postings_full <-left_join(all_dates,new_postings, by = c("date" = "date_published"))

new_postings_full %<>% 
  ungroup() %>% 
  mutate(recipe_author = ifelse(is.na(recipe_author),0,1)) %>% 
  group_by(date) %>% 
  summarise(total_new_authors = sum(recipe_author)) %>% 
  ungroup() %>% 
  mutate(cumulative_new_authors = cumsum(total_new_authors))

new_postings_full  %<>% 
  mutate(trend = row_number()) %>% 
  mutate(pub_year = year(date))

new_postings_full_split <- split(new_postings_full,new_postings_full$pub_year)
trendy <- function(split_data){
  # split_data <- test_split[[1]]
  lm_trend <- lm(cumulative_new_authors~trend,data = split_data)
  split_data$predicted_values <- predict(lm_trend, newdata = split_data)
  split_data$rate <- lm_trend$coefficients["trend"]
  split_data
}

new_postings_full <- map(new_postings_full_split, trendy) %>% bind_rows()
new_postings_full%<>% filter(predicted_values > 0) %>% rename(Rate = rate)

plot_ly(data = date_summary_full, y=~predicted_values, x = ~date, 
        type = 'scatter', mode = 'markers', color = ~Rate, showlegend=FALSE) %>% 
  layout(
    xaxis = list(range = c(min(date_summary_full$date)-5,max(date_summary_full$date) + 5),
                 title = ""),
    yaxis = list(title = "Number of New Authors"))

```

# Single/Multiple Contributors
As the data has shown so far, new authors are very important to increasing the number of recipes 
added to the community cake collection.  The data has also shown that multiple contributions from 
a single author can have a large impact.  The difference in contributions are shown below. Those 
that have contributed more than one recipe are classifed as as 'multiple contributors' and those 
that have contributed one recipe as 'single contributors'.  The plot shows in recent years 
there is a widening gap in contributions between the two groups.  This is in part 
due to a decreased rate of new authorship, but there appears to be more than a simple relationship 
in the data.

```{r, echo=F, message=F}
multiple_cont <- recipe_info %>% 
  group_by(recipe_author) %>% 
  summarise(count = n()) %>% 
  filter(count > 1) %>% 
  mutate(multiple_contributions = T)
#only 361 (31%) have contributed multiple cakes
#what percent of new cakes contributed are contributed by first time vs repeat 

recipe_info_multi <- left_join(recipe_info,multiple_cont) %>% 
  mutate(multiple_contributions = ifelse(is.na(multiple_contributions),
                                         "single_contribution",
                                         "multiple_contributions")) %>% 
  mutate(pub_year = year(date_published)) %>%
  mutate(pub_month = month(date_published)) %>% 
  group_by(pub_year, pub_month,multiple_contributions) %>% 
  summarise(count_recipes = n()) %>% 
  ungroup() %>% 
  spread(multiple_contributions,count_recipes) %>% 
  arrange(pub_year,pub_month) %>% 
  mutate(m_y_order = row_number()) %>% 
  mutate(year_month = paste0(pub_year,"_",pub_month)) 
recipe_info_multi[is.na(recipe_info_multi)] <- 0
  
recipe_info_multi %<>%
  mutate(total_contributions = multiple_contributions + single_contribution) %>% 
  mutate(percent_multiple_contributions = multiple_contributions/total_contributions) %>% 
  mutate(percent_single_contributions = single_contribution/total_contributions) %>% 
  mutate(year_month_date = as.Date(paste0(year_month,"_01"),f = "%Y_%m_%d"))
  
#Over time, the difference between the contributions of single and multiple contributors have 
# diverged (test).  Contests drive increased numbers of recipes from both groups, but multiple 
# contributors seem to be driven more (test).
plot_ly(data = recipe_info_multi,
        # x = ~forcats::fct_reorder(year_month,m_y_order),
        x = ~year_month_date,
        y = ~multiple_contributions,
        type = 'scatter', mode = 'lines', name = "multiple") %>% 
  add_trace(y = ~single_contribution, name = "single") %>% 
  layout(
    xaxis = list(title = ""),
    yaxis = list(title = "Recipes Added"))
```

## Model the differences

A model based approach lets us see the more nuanced differences between the two groups of 
contributors.  Many possible variables were tested to analyze the difference in the groups, but the 
most explanatory variables are time and the presence of an active contest.  

On average, both single 
and multiple contributor groups are expected to contribute more recipes over time, but at different
rates. The model projects recipe contributions to increase by one every 4 months for multiple
contributors and  one every 14 months for single contributors. 

Active contests were also found to increase the expected amount of new recipes added by both groups.
The impact of contests were found by investigating the spikes found at semi regular points on the 
above graph.  An example can be seen in November of 2014 with the 'Your best Breakfast Baked Good'
contest.  Both groups saw a spike in contributions above what would normally be expected.  
Active contests were found to be dependent on the type of contributor. The monthly expected recipe
count increases by 16 for multiple contributors and by 10 for single contributors

```{r, echo=F, message = F}
recipe_info_multi_lm <- recipe_info_multi %>% 
  select(m_y_order,year_month,multiple_contributions, single_contribution) %>% 
  gather(contributor_type, amount_contributed, -m_y_order,-year_month)
  
test_lm <- lm(amount_contributed~m_y_order*contributor_type, data = recipe_info_multi_lm)
#This shows that the number of recipes commited has increased on average by 1 every 5 months, 
#this rate is slower for single contributors.
# summary(test_lm)
#data based on researching spikes in both contributor types.  Found that many of the spikes in 
#contributions were associated to a theme ie carrot or cardmom cakes.  These associations map well
#to active contests during the month observed.
contests <- data_frame(year_month = c("2009_9","2010_1","2009_6","2013_5","2011_2",
                                      "2014_11","2013_10","2010_11","2011_3","2011_6","2014_7"),
                       
                       contest_title = c("Best Apple Cake","Best Chocolate Cake", "Best Layer Cake",
                                        "Best Recipe with Coconut",
                                         "Best Recipe with Cardmom",
                                         "Your Best Breakfast Baked Good","Best Recipe for Frosting",
                                         "Your Best Non-Pie Thanksgiving Dessert",
                                         "Your Best Carrot Recipe",
                                         "Your Best Gluten-Free Baked Good",
                                         "Your Best Berry Recipe"
                                         )
                       )

recipe_info_multi_lm %<>% 
  left_join(contests) %>% 
  mutate(active_contest = ifelse(is.na(contest_title),"no","yes"))

test_lm <- lm(amount_contributed~m_y_order + contributor_type + active_contest +
              m_y_order*contributor_type + contributor_type*active_contest,
              data = recipe_info_multi_lm)

# this model projects recipe contributions to increase by one every 4 months for multiple
# contributors and  one every 14 months for single contributors.  Active contests increase the 
# monthly expected recipe count by 16 for multiple contributors and by 10 for single contributors
summary(test_lm)
```

#Active Multiple contributors

As shown above, on average multiple contributors add more recipes per month than single contributors 
and they are expected to continue to do so.  By summing the total number of recipes contributed by
group, multiple contributors have authored 67% of the cake recipes in this data while only 
comprising 31% of the population of authors.

With this in mind, lets look at the survivorship of multiple contributors.  

For this analysis, an 'active conttributor' is a multiple contributor who has published in the last
135 days.  135 days was chosen because it is the 75% percentile of days between publishing for
multiple contributors.  This means that 75% of the intra publishing times for new recipes is 
less than 135 days.  

By totalling the number of active contributors over every day within the data, we get the plot 
below.  The plot below shows a decrease in the number of active multiple contributors in recent 
history.  

```{r, echo = F, message=F}
multi_info <- inner_join(recipe_info, multiple_cont) %>% 
  group_by(recipe_author) %>% 
  arrange(date_published) %>% 
  mutate(time_since_last_post = difftime(date_published,lag(date_published),units = "days" )) %>% 
  ungroup() %>% 
  mutate(time_since_last_post = as.numeric(time_since_last_post)) 

# multi_info %>% 
#   filter(!is.na(time_since_last_post)) %>% 
#   .$time_since_last_post %>% 
#   quantile()
#75% of times between posts are 135, this will serve as a cutoff for 'active repeat user'

author_split <- split(multi_info, multi_info$recipe_author)

# auth_split <- author_split[["aargersi"]]
# time_liner(days_since_cake = as$time_since_last_post,churn_time = 135)
# days_since_cake %<>% as.numeric() 
# days_since_cake <- as$time_since_last_post
# churn_time <- 135
length_period <- function(dates, churn_time){
  if(length(dates) == 1){
    churn_time
  }else{
    as.numeric(difftime(max(dates),min(dates),units = "days") )+ 135
  }
}

time_liner <- function(auth_split, churn_time){
  days_since_cake <- auth_split$time_since_last_post
  days_since_cake[is.na(days_since_cake)] <- 0
  period <- rep(NA, length(days_since_cake))
  per <- 1
  while(length(days_since_cake)){
    
    if(length(days_since_cake) == 1){
      period[is.na(period)] <- per
      days_since_cake <- NULL
    }else{
      days_since_cake[1] <- 0
      if(max(days_since_cake) < 135){
        period[is.na(period)] <- per
        days_since_cake <- NULL
      }else{
      first_new_period <- min(which(days_since_cake >= churn_time))
      within_window <- days_since_cake < churn_time
      within_window[first_new_period:length(within_window)] <- F
      period[c(rep(F,sum(!is.na(period))),within_window)] <- per
      days_since_cake <- days_since_cake[!within_window]
      per <- per + 1
      }
    }
    
    
  }
  auth_split$period <- period
  auth_split %>% 
    group_by(period) %>% 
    mutate(period_length = length_period(date_published, churn_time)) %>% 
    ungroup()
}

# time_liner(auth_split = author_split[["adinab"]], 135)

multi_info <- map(author_split,time_liner,churn_time = 135) %>% bind_rows()

multi_info_summary <- multi_info %>% 
  group_by(recipe_author,period) %>% 
  summarise(period_begin = min(date_published),
            number_contributions_period = n(),
            period_length = max(period_length)) %>% 
  ungroup()

really_now_single <- multi_info_summary %>% 
  group_by(recipe_author) %>% 
  filter(period == max(period)) %>% 
  filter(number_contributions_period == 1) %>% 
  ungroup() %>% 
  filter(period_begin + days(period_length) < max(recipe_info$date_published))

multi_info_summary <- anti_join(multi_info_summary, really_now_single, by = c("recipe_author" = "recipe_author", "period" = "period"))

multi_info_summary %<>% 
  mutate(auth_period = paste0(recipe_author,"_", period))
multi_info_summary_split <- split(multi_info_summary,multi_info_summary$auth_period)
test <- multi_info_summary_split[[1]]
date_expander <- function(split_data){
data_frame(
  auth_period = split_data$auth_period,
  contributions = split_data$number_contributions_period,
  dates = seq.Date(split_data$period_begin,split_data$period_begin + 
                     days(split_data$period_length),by = "days")
)
}
testo <- map(multi_info_summary_split, date_expander) %>% bind_rows() 
testo_1 <- testo %>%
  filter(dates < max(recipe_info$date_published)) %>% 
  group_by(dates) %>% 
  summarise(count_periods = n())

plot_ly(data = testo_1, y = ~count_periods, x = ~dates,
       type = 'scatter', mode = 'markers' )
```

# Summary

The number of people contributing to the community based collection of cake recipes on 
Food52 is not as large as it has been, but because of the increased rate of per author publications
the realative number of new recipes the commuity enjoys each month has remained high.  To ensure
the continued vigor in the community, Food52 should continue to do all it can to increase enagement 
of multiple contributors and new contributors alike.  Well publicized competitions and growing
the vibrant community of users are proven steps to do this. 


